Exploring and Predicting House Price in Nanjing: Geographically Imbalanced Distribution Across Districts and the Crucial Role of Unit Price and Bedroom Count in Prediction

### 2023-12-13 output/paper/paper.qmd
Update references

---
title: "Exploring and Predicting House Price in Nanjing: Geographically Imbalanced Distribution Across Districts and the Crucial Role of Unit Price and Bedroom Count in Predictioin"
author: "Yufei Liu"
thanks: "Code and data are available at: https://github.com/Florence-Liu/house-price"
date: today
date-format: long
abstract: "House prices in Nanjing have increased rapidly and became a concern in recent years. This paper examines the relationship between house price in Nanjing, structural attributes of the property, and location using open data collected from Lianjia.com. We then predict house price in Nanjing with different models. We find a geographically-imbalanced distribution of house price in Nanjing with large variance within and between each district. By comparing the MAE, RMSE, and $R^2$, we find the Random Forest model has a better prediction performance than the Multiple Linear Regression model. Unit price and number of bedrooms in the house tend to be importance features for predicting house price in Nanjing. Further work could use spatial data to include the spatial effect in the model, and tune hyperparameters to improve model performance."
format: pdf
number-sections: true
toc: false
bibliography: references.bib
---

@@

In the process of RF, the number of trees in the forest, the number of candidate variables at each split, and the minimum size of terminal nodes are three important parameters that affect the performance of the RF model [@citeHyper]. A common choice of the number of candidate variable $m$ is $m\approx \sqrt{p}$ for classification trees, and $m=\frac{p}{3}$ for regression trees [@citeM]. We will use `randomForest` function in package `randomForest` in `R` to fit the RF model. The function by default set the minimum size of nodes $n_{min}$ to be 5, and number of trees $\text{ntree}$ to be 500. The number of candidate variables $m$ is determined by $m=\frac{p}{3}$ for regression trees as mentioned before. We will assess the importance of each predictor variable in the RF model by measuring either the percentage increase in Mean Square Error (MSE) or the increase in Node Purity for each split in trees.

@@

## Random Forest {#sec-result_mlg}

@fig-model shows the importance of variable of the random forest model based on two metrics, Mean Square Error (MSE), and Node Purity. The percentage increase in MSE ($\%\text{IncMSE}$) is calculated by how much in percentage MSE increases without the predictor. The higher value indicates more important features for making predictions. The increase in Node Purity ($\text{IncNodePurity}$) measure by how much node purity increases when splitting on a specific feature. It is usually calculated by training RSS, and same as $\%\text{IncMSE}$, the higher value indicates higher influence on prediction performance. From @fig-model we find that the three most important predictors are the same based on $\%\text{IncMSE}$ and $\text{IncNodePurity}$, which are `Unit_Price`, `Area`, and `Bedroom`. The two least important predictors are also the same, which are `Furnished` and `Facing_South`.

@@

There are three metrics being used to evaluate the two models: MAE, RMSE, and $\mbox{R}^2$. Mean Absolute Error (MAE) measures the average absolute difference between the predicted value and actual values [@citeMAE]. Root Mean Squared Error (RMSE) measures the square root of the average squared difference between the predicted value and the actual value [@citeRMSE]. The coefficient of determination ($R^2$) measures the proportion of the total variability in the response variable that can be explained by the model [@citeR2]. The expressions for the three metrics are shown in @tbl-expression. Lower values in MAE and RMSE and higher value (between 0 and 1) in $R^2$ suggest better model performance.

@@

R2 <- c(r2_mlg_all, r2_mlg_train, r2_mlg_test, r2_rf_all, r2_rf_train, r2_rf_test)
MAE <- c(mae_mlg_all, mae_mlg_train, mae_mlg_test, mae_rf_all, mae_rf_train, mae_rf_test)
RMSE <- c(rmse_mlg_all, rmse_mlg_train, rmse_mlg_test, rmse_rf_all, rmse_rf_train, rmse_rf_test)
Model <- c("MLG all", "MLG train", "MLG test", "RF all", "RF train", "RF test")
metrics <- data.frame(Model, MAE, RMSE, R2)
kable(metrics, booktabs = TRUE, linesep = "", 
      col.names = c("Model", "MAE", "RMSE", "$\\mbox{R}^2$"), escape = F)

@@

The economical development in each district also contributes the difference in house price across districts. There are two main commercial districts in Nanjing, one in the city center in Gulou district and one in the CBD in Jianye district. Due to limited available space and large population density, high-rise condominiums and luxury apartments are two main types of housing in Gulou and Jianye. These two types of house usually demand higher price due to the modern amenities and location. In addition, commercial districts are accompanied by more job opportunities and higher income level, which attracts those who are seeking more convenient living environment and houses closer to companies to shorten commuting time. The high demand for housing in commercial districts also affects the rental market, thus increasing investment enthusiasm. For Gulou district, it has a third type of house, old apartment built early in 1980s. Since it was developed early, the amenities sometimes function improperly, and management is lax and inefficient with small floor areas. However, such type of house still has a high total price. This is related to its surrounding facilities especially schools. In Nanjing, public schools in compulsory education period that a child may attend is determined by the household registration system, that is primary school and middle school. As a result, parents who wish their children to attend a better public school may choose to buy an apartment within the school district[@citeEdu]. Since it was developed early, many qualified and renowned public schools are located in Gulou district, resulting in the high demand of houses in school district divisions and thus higher house price.

We could also notice the highest house price on sale in Nanjing was about 40,000,000 yuan, which is relatively not very high compared with other economically-developed cities such as Shanghai. One possible reason is that some houses are not listed and sold publicly. Private sales are common among luxurious houses for which price could be over hundreds of million yuan. Another possible reason is government regulation and economic recession. The Chinese government has implemented several policies including land policies, fiscal policies, and monetary policies in various cities to cool down the heated market and housing speculations after 2021 [@citePolicy]. These policies had successfully cooled down the market, however, in an unexpected speed. The sudden policy change and higher unemployment rate have made buying houses a prohibitive choice, and struck the real estate developers who had not been recovered completely from pandemic. Although in 2023, the Nanjing government had announced several promoting policies to help revive the market including decreased interest rate and cancellation of preconditions for buying houses in inner city, the affect was limited. Consumers took a low expectation on current market, greatly reduced investments and speculations on houses. As a result, the demand of house has been decreased compared with 2018, thus affecting the house price dynamics.

@@

For models we used in the analysis, we could also consider some improvements in future work. The validity of MLR model was challenged by the potential multicolinearity concern as shown in @fig-cor, which affected the reliability of the estimates of coefficients. Additionally, although 
the difference between performance on training data and testing data for two models was not significant, we should still be careful about overfitting. The RF model is less prone to overfitting by ensemble learning algorithm, however, a complex tree structure may also arise the problem of overfitting. In this case, we may consider using shrinkage methods such as Lasso or Ridge regression to mitigate the problem of multicolinearity and overfitting for the MLR model. For the RF model, we did not tune hyperparameters in the model, that is the number of tree, the number of candidate predictors at each split, and the minimum size of nodes. Although the model yielded the prediction results with a relatively low MAE and RMSE and a high $R^2$, we could still enhance the model performance by tuning hyperparameter. It will also increase the robustness and generalizability of the model and ensure the optimal performance.

>>>>

---
title: "Exploring and Predicting House Price in Nanjing: Geographically Imbalanced Distribution Across Districts and the Crucial Role of Unit Price and Bedroom Count in Prediction"
author: "Yufei Liu"
thanks: "Code and data are available at: https://github.com/Florence-Liu/house-price"
date: today
date-format: long
abstract: "House prices in Nanjing have increased rapidly and became a concern in recent years. This paper examines the relationship between house price in Nanjing, structural attributes of the property, and location using open data collected from Lianjia.com. We then predict house price in Nanjing with different models. We find a geographically-imbalanced distribution of house price in Nanjing with large variance within and between each district. By comparing the MAE, RMSE, and $R^2$, we find the Random Forest model has a better prediction performance than the Multiple Linear Regression model. Unit price and number of bedrooms in the house tend to be importance features for predicting house price in Nanjing. Further work could use spatial data to include the spatial effect in the model, and tune hyperparameters to improve model performance."
format: pdf
number-sections: true
toc: false
bibliography: references.bib
---

@@

In the process of RF, the number of trees in the forest, the number of candidate variables at each split, and the minimum size of terminal nodes are three important parameters that affect the performance of the RF model. A common choice of the number of candidate variable $m$ is $m\approx \sqrt{p}$ for classification trees, and $m=\frac{p}{3}$ for regression trees [@citeM]. We will use `randomForest` function in package `randomForest` in `R` to fit the RF model. The function by default set the minimum size of nodes $n_{min}$ to be 5, and number of trees $\text{ntree}$ to be 500. The number of candidate variables $m$ is determined by $m=\frac{p}{3}$ for regression trees as mentioned before. We will assess the importance of each predictor variable in the RF model by measuring either the percentage increase in Mean Square Error (MSE) or the increase in Node Purity for each split in trees.

@@

## Random Forest {#sec-result_rf}

@fig-model shows the importance of variable of the random forest model based on two metrics, Mean Square Error (MSE), and Node Purity. The percentage increase in MSE ($\%\text{IncMSE}$) is calculated by how much in percentage MSE increases without the predictor. The higher value indicates more important features for making predictions. The increase in Node Purity ($\text{IncNodePurity}$) measure by how much node purity increases when splitting on a specific feature. It is usually calculated by training RSS, and same as $\%\text{IncMSE}$, the higher value indicates higher influence on prediction performance [@citeVarImp]. From @fig-model we find that the three most important predictors are the same based on $\%\text{IncMSE}$ and $\text{IncNodePurity}$, which are `Unit_Price`, `Area`, and `Bedroom`. The two least important predictors are also the same, which are `Furnished` and `Facing_South`.

@@

There are three metrics being used to evaluate the two models: MAE, RMSE, and $\mbox{R}^2$ [@citeMetrics]. Mean Absolute Error (MAE) measures the average absolute difference between the predicted value and actual values. Root Mean Squared Error (RMSE) measures the square root of the average squared difference between the predicted value and the actual value. The coefficient of determination ($R^2$) measures the proportion of the total variability in the response variable that can be explained by the model. The expressions for the three metrics are shown in @tbl-expression. Lower values in MAE and RMSE and higher value (between 0 and 1) in $R^2$ suggest better model performance.

@@

R2 <- c(r2_mlg_all, r2_mlg_train, r2_mlg_test, r2_rf_all, r2_rf_train, r2_rf_test)
MAE <- c(mae_mlg_all, mae_mlg_train, mae_mlg_test, mae_rf_all, mae_rf_train, mae_rf_test)
RMSE <- c(rmse_mlg_all, rmse_mlg_train, rmse_mlg_test, rmse_rf_all, rmse_rf_train, rmse_rf_test)
Model <- c("MLR all", "MLR train", "MLR test", "RF all", "RF train", "RF test")
metrics <- data.frame(Model, MAE, RMSE, R2)
kable(metrics, booktabs = TRUE, linesep = "", 
      col.names = c("Model", "MAE", "RMSE", "$\\mbox{R}^2$"), escape = F)

@@

The economical development in each district also contributes the difference in house price across districts. There are two main commercial districts in Nanjing, one in the city center in Gulou district and one in the CBD in Jianye district. Due to limited available space and large population density, high-rise condominiums and luxury apartments are two main types of housing in Gulou and Jianye. These two types of house usually demand higher price due to the modern amenities and location. In addition, commercial districts are accompanied by more job opportunities and higher income level, which attracts those who are seeking more convenient living environment and houses closer to companies to shorten commuting time. The high demand for housing in commercial districts also affects the rental market, thus increasing investment enthusiasm. For Gulou district, it has a third type of house, old apartment built early in 1980s. Since it was developed early, the amenities sometimes function improperly, and management is lax and inefficient with small floor areas. However, such type of house still has a high total price. This is related to its surrounding facilities especially schools. In Nanjing, public schools in compulsory education period that a child may attend is determined by the household registration system, that is primary school and middle school [@citeEdu]. As a result, parents who wish their children to attend a better public school may choose to buy an apartment within the school district[@citeParent]. Since it was developed early, many qualified and renowned public schools are located in Gulou district, resulting in the high demand of houses in school district divisions and thus higher house price.

We could also notice the highest house price on sale in Nanjing was about 40,000,000 yuan, which is relatively not very high compared with other economically-developed cities such as Shanghai. One possible reason is that some houses are not listed and sold publicly. Private sales are common among luxurious houses for which price could be over hundreds of million yuan. Another possible reason is government regulation and economic recession. The Chinese government has implemented several policies including land policies, fiscal policies, and monetary policies in various cities to cool down the heated market and housing speculations after 2021 [@citePolicy]. These policies had successfully cooled down the market, however, in an unexpected speed. The sudden policy change and higher unemployment rate have made buying houses a prohibitive choice, and struck the real estate developers who had not been recovered completely from pandemic [@citeAffect]. Although in 2023, the Nanjing government had announced several promoting policies to help revive the market including decreased interest rate and cancellation of preconditions for buying houses in inner city, the affect was limited [@citeNJPolicy]. Consumers took a low expectation on current market, greatly reduced investments and speculations on houses. As a result, the demand of house has been decreased compared with 2018, thus affecting the house price dynamics.

@@

For models we used in the analysis, we could also consider some improvements in future work. The validity of MLR model was challenged by the potential multicolinearity concern as shown in @fig-cor, which affected the reliability of the estimates of coefficients. Additionally, although 
the difference between performance on training data and testing data for two models was not significant, we should still be careful about overfitting. The RF model is less prone to overfitting by ensemble learning algorithm, however, a complex tree structure may also arise the problem of overfitting. In this case, we may consider using shrinkage methods such as Lasso or Ridge regression to mitigate the problem of multicolinearity and overfitting for the MLR model. For the RF model, we did not tune hyperparameters in the model, that is the number of tree, the number of candidate predictors at each split, and the minimum size of nodes. Although the model yielded the prediction results with a relatively low MAE and RMSE and a high $R^2$, we could still enhance the model performance by tuning hyperparameters. It will also increase the robustness and generalization of the model and ensure the optimal performance.
